Numerical experiment investigating the influence of drag on trajectory patterns of floating macroalgae

### 2021-03-24 thesis_chapter_3.Rmd
updates

@@

dis_plot <- ggplot(data = total_data, aes(x = type, y = distance, show.legend = FALSE))+
  geom_boxplot(data = p_RK4, notch = TRUE, notchwidth = 1, show.legend = FALSE, stat = "boxplot") +
  geom_boxplot(notch = TRUE, notchwidth = 1, show.legend = FALSE, stat = "boxplot") +
  #scale_x_discrete(limits = pos)+
  stat_compare_means(method = "kruskal.test", label = "p.signif", label.x.npc = c("right"))+
  theme_bw()+
  ylab("Distance (km)")+
  xlab("Scenario")+
  #coord_flip() +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

custom_labs <- c("1","2", "3", "4", "5", "7", "8", "9", "10", "11", "12", "13") #create custom labels and add them to dis_plot

@@

```{r distance plot, message=FALSE, warning=FALSE, echo = FALSE, fig.cap="Figure 4: Boxplots of total distance travelled by all particles from the release site over the course of the simulation for scenarios 1 - 13, for January 2018 (summer). The x-axis represents the scenario number and the y-axis is the distance travelled in kilometers. The lower and upper hinges correspond to the first and third quartiles. The whiskers represent the range, solid black lines represent the median and black dots are outliers. The reader is referred to tables 1 and 2 for details pertaining to each scenario."}
dis_plot
```

@@

kelp_acc$type <- "Kelp" # add extra column 'type' to dataframe and assigned 'kelp' label
kelp_acc_mean$type <- "Kelp" # add extra column 'type' to dataframe and assigned 'kelp' label

@@

pass_acc$type <- "Passive" # add extra column 'type' to dataframe and assigned 'kelp' label
pass_acc_mean$type <- "Passive" # add extra column 'type' to dataframe and assigned 'kelp' label

@@

point_kelp_acc <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  geom_point(data = kelp_acc, aes(x = lon, y = lat, colour = age)) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
  xlab("Longtitude") +
  ylab("Latitude") +
  ggtitle("") +
  guides(fill=guide_legend(title="Age")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

point_kelp_acc <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  geom_point(data = kelp_acc, aes(x = lon, y = lat, colour = age)) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
  xlab("Longtitude") +
  ylab("Latitude") +
  ggtitle("") +
  #scale_color_viridis_c(name = "Age") +
  guides(fill=guide_legend(title="Age")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

@@

ggarrange(plot_A, plot_B, labels = c("AUTO"), ncol = 1, nrow = 2)

@@

The ocean is comprised of time-varying flows and changes in an objects velocity that can cause the object to be exposed to different flow patterns over time. The lack of dissimilarity in trajectory between different cross-sectional area types (minimum, mean and maximum) suggests that reduction in velocity in relation to cross-sectional area is negligible. A similar result was found by @gouvello2020 who investigated the effects of swimming behavior on sea turtle hatchling dispersal in the Agulhas region. The results suggested that the hatchling trajectories are mostly influenced by the ocean currents in the first year of hatching due to  as juveniles do are not swim fast enough to alter their overall trajectory. Therefore, the differences in cross-sectional area types may not be significant enough to have an effect on the relevant zonal and meridional velocity vectors. Another possible reason is the resolution of the ocean model used. The resolution of the underlying ocean model is an important aspect of Lagrangian ocean modeling and must be able to resolve sub-grid scale ocean processes for finer scale applications. A study by @hart2018 assessed the inclusion of stochastic motion, wind and currents into forecasting for search and rescue using the same model ocean model used in this study. The authors found that the inclusion of Brownian motion greatly increased the accuracy in representing sub-grid scale processes and that the inclusion of wind, currents and stochastic motion greatly improved forecasting applications. Therefore, the authors argue that the lack of dissimilarity between simulations of different cross-sectional area and drag exposure scenarios is not as a result of the resolution of the model.

The same is also true when comparing different hydrodynamic and wind drag exposure scenarios. These scenarios reflect different buoyancy situations which result in different amounts of surface area exposed to hydrodynamic and/or wind drag. The similarities between the different combined drag scenarios suggest that higher magnitudes of wind exposure do not significantly alter trajectories. Instead, the results from this study suggest that the inclusion of drag forces in simulating macroalgal trajectory may result in improving accuracy of entrainment patterns. All the simulations that included any form of drag resulted in a  high density of particles entrained in a eddy. The entrainment and expulsion from eddies and vortices are an important characteristic of floating macroalgae. Accuracy in predicting entrainment and expulsion of macroalgae-rafts from eddies has been investigated previously by @putman2018 and @putman2020, who show that including windage greatly improves entrainment patterns when comparing virtual and real Sargassum rafts. However, these studies investigated large rafts and not floating individuals. Inertia is based on the size/mass of the raft, while hydrodynamic and wind drag are based on current velocity and surface area. Therefore, the role of inertia may be negligible for a solitary individual and rather other forms of drag, such as in this study, should be considered. The results from this study suggest that the inclusion of hydrodynamic and/or wind drag is an important component when simulating virtual kelp particles, and will greatly increase the accuracy of entrainment and expulsion from sub-grid mesoscale oceanographic features, such as eddies and vortices.

>>>>

data_dis <- total_data %>% 
  select(!(kelpvelU)) %>%
  select(!(kelpvelV)) %>%
  select(!(velocity))

total_dis_data <- rbind(p_RK4, data_dis)

@@

dis_plot <- ggplot(data = total_dis_data, aes(x = type, y = distance, show.legend = FALSE))+
  geom_boxplot(notch = TRUE, notchwidth = 1, show.legend = FALSE, stat = "boxplot") +
  #scale_x_discrete(limits = pos)+
  stat_compare_means(method = "kruskal.test", label = "p.signif", label.x.npc = c("right"))+
  theme_bw()+
  ylab("Distance (km)")+
  xlab("Scenario")+
  #coord_flip() +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

custom_labs <- c("1","2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12", "13") #create custom labels and add them to dis_plot

@@

```{r distance plot, message=FALSE, warning=FALSE, echo = FALSE, fig.cap="Figure 4: Boxplots of total distance travelled by all particles from the release site over the course of the simulation for scenarios 1 - 13, for January 2018 (summer). The x-axis represents the scenario number and the y-axis is the distance travelled in kilometers. The lower and upper hinges correspond to the first and third quartiles. The whiskers represent the range, solid black lines represent the median and black dots are outliers. The reader is referred to tables 1 and 2 for details pertaining to each scenario."}

dis_plot
ggsave(filename = "figure_4.jpg", plot = last_plot(), device = "jpeg", path = "figures/chapter_3/", width = 8, height = 5, dpi = 600)

@@

kelp_acc$type <- "Kelp" # add extra column 'type' to dataframe and assigned 'kelp' label
kelp_acc_mean$type <- "Kelp" # add extra column 'type' to dataframe and assigned 'kelp' label
kelp_acc$age <- with(kelp_acc, age/3600)

@@

pass_acc$type <- "Passive" # add extra column 'type' to dataframe and assigned 'kelp' label
pass_acc_mean$type <- "Passive" # add extra column 'type' to dataframe and assigned 'kelp' label
pass_acc$age <- with(pass_acc, age/3600)

@@

point_kelp_acc <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  geom_point(data = kelp_acc, aes(x = lon, y = lat, colour = age)) +
  #scale_fill_continuous(type = "gradient", limits = c(0, 30)) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) +
  xlab("Longtitude") +
  ylab("Latitude") +
  ggtitle("") +
  guides(fill=guide_legend(title="Age")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

@@

ggarrange(plot_A, plot_B, labels = c("AUTO"), ncol = 1, nrow = 2)
ggsave(filename = "figure_5.jpg", plot = last_plot(), device = "jpeg", path = "figures/chapter_3/", width = 7, height = 7, dpi = 600)

@@

The ocean is comprised of time-varying flows and changes in an objects velocity that can cause the object to be exposed to different flow patterns over time. The lack of dissimilarity in trajectory between different cross-sectional area types (minimum, mean and maximum) suggests that reduction in velocity in relation to cross-sectional area is negligible. A similar result was found by @gouvello2020 who investigated the effects of swimming behavior on sea turtle hatchling dispersal in the Agulhas region. The results suggested that the hatchling trajectories are mostly influenced by the ocean currents in the first year of hatching due as juveniles do not swim fast enough to alter their overall trajectory. Therefore, the differences in cross-sectional area types may not be significant enough to have an effect on the relevant zonal and meridional velocity vectors. Another possible reason is the resolution of the ocean model used. The resolution of the underlying ocean model is an important aspect of Lagrangian ocean modeling and must be able to resolve sub-grid scale ocean processes for finer scale applications. A study by @hart2018 assessed the inclusion of stochastic motion, wind and currents into forecasting for search and rescue using the same model ocean model used in this study. The authors found that the inclusion of Brownian motion greatly increased the accuracy in representing sub-grid scale processes and that the inclusion of wind, currents and stochastic motion greatly improved forecasting applications. Therefore, the authors argue that the lack of dissimilarity between simulations of different cross-sectional area and drag exposure scenarios is not as a result of the resolution of the model.

The same is also true when comparing different hydrodynamic and wind drag exposure scenarios. These scenarios reflect different buoyancy situations which result in different amounts of surface area exposed to hydrodynamic and/or wind drag. The similarities between the different combined drag scenarios suggest that higher magnitudes of wind exposure do not significantly alter trajectories. Instead, the results from this study suggest that the inclusion of drag forces in simulating macroalgal trajectory may result in improving accuracy of entrainment patterns. All the simulations that included any form of drag resulted in a high density of particles entrained in an eddy. The entrainment and expulsion from eddies and vortices are an important characteristic of floating macroalgae. Accuracy in predicting entrainment and expulsion of macroalgae-rafts from eddies has been investigated previously by @putman2018 and @putman2020, who show that including windage greatly improves entrainment patterns when comparing virtual and real Sargassum rafts. However, these studies investigated large rafts and not floating individuals. Inertia is based on the size/mass of the raft, while hydrodynamic and wind drag are based on current velocity and surface area. Therefore, the role of inertia may be negligible for a solitary individual and rather other forms of drag, such as in this study, should be considered. The results from this study suggest that the inclusion of hydrodynamic and/or wind drag is an important component when simulating virtual kelp particles, and will greatly increase the accuracy of entrainment and expulsion from sub-grid mesoscale oceanographic features, such as eddies and vortices.
